---
title: "Cost Of Voting -Gerber"
author: "Soryn Lurding"
date: "October 3, 2024"
output: pdf_document
---

## Learning Objectives

In this assignment, you will practice your T-Test and R programming skills. You will use a fixed effects model and perform a difference-in-differences analysis. 
Please answer the questions in a R markdown file and “Knit” the file so that I can see your analysis.

## Experiments 

Reanalysis of Gerber, Green and Larimer (2008)
‘Why do large numbers of people vote, despite the fact that, as Hegel once observed, “the casting of a single vote is of no significance where there is a multitude of electors”?’

This is the question that drives the experimental analysis of Gerber, Green and Larimer (2008). If it is irrational to vote because the costs of doings so (time spent informing oneself, time spent getting to the polling station, etc) are clearly greater than the gains to be made from voting (the probability that any individual voter will be decisive in an election are vanishingly small), then why do we observe millions of people voting in elections? One commonly proposed answer is that voters may have some sense of civic duty which drives them to the polls. Gerber, Green and Larimer investigate this idea empirically by priming voters to think about civic duty while also varying the amount of social pressure voters are subject to.
In a field experiment in advance of the 2006 primary election in Michigan, nearly 350,000 voters were assigned at random to one of four treatment groups, where voters received mailouts which encouraged them to vote, or a control group where voters received no mailout. The treatment and control conditions were as follows:

	Treatment 1 (“Civic duty”): Voters receive mailout reminding them that voting is a civic duty
	Treatment 2 (“Hawthorne”): Voters receive mailout telling them that researchers would be studying their turnout based on public records
	Treatment 3 (“Self”): Voters receive mailout displaying the record of turnout for their household in prior elections.
	Treatment 4 (“Neighbors”): Voters receive mailout displaying the record of turnout for their household and their neighbours’ households in prior elections.
	Control: Voters receive no mailout.
	
Load the replication data for Gerber, Green and Larimer (2008). This data is stored in a .Rdata format, which is the main way to save data in R. Therefore you will not be able to use read.csv but instead should use the function load.
```{r}
# You will need to change the file location for the code to work.
load("~/Downloads/gerber_green_larimer.Rdata")
```

Once you have loaded the data, familiarise yourself with the the gerber object which should be in your current envionment. Use the str and summary functions to get an idea of what is in the data. There are 5 variables in this data.frame:
Variable name	Description

- voted:	Indicator for whether the voter voted in the 2006 election (1) or did not vote (0)
- treatment:	Factor variable indicating which treatment arm (or control group) the voter was allocated to
- sex:	Sex of the respondent
- yob:	Year of birth of the respondent
- p2004: Indicator for whether the voter voted in the 2004 election (Yes) or not (No)

1. Calculate the turnout rates, "voted", for each of the experimental groups (4 treatments, 1 control). Calculate the number of individuals allocated to each group. Recreate table 2 on p. 38 of the paper.
```{r}
library(tidyverse)
library(modelsummary)
datasummary((`PCT Voted`=voted) ~ Mean*treatment, data=gerber)
```

2. Use the following code to create three new variables in the data.frame. First, a variable that is equal to 1 if a respondent is female, and 0 otherwise. Second, a variable that measures the age of each voter in years at the time of the experiment (which was conducted in 2006). Third, a variable that is equal to 1 if the voter voted in the November 2004 Midterm election.
```{r}
## Female dummy variable
gerber$female <- ifelse(gerber$sex == "female", 1, 0)

## Age variable
gerber$age <- 2006 - gerber$yob

## 2004 variable
gerber$turnout04 <- ifelse(gerber$p2004 == "Yes", 1, 0)
```
3. Using these variables, conduct balance checks to establish whether there are potentially confounding differences between treatment and control groups. You do this by using the variables female, age, and turnout04 as dependent variables. Use just the factor variable of treatment as your explanatory variable. Can you conclude from the results that randomization worked? How do you know?

```{r}
exp1 <-lm(female ~ factor(treatment), data =gerber)
exp2 <-lm(age ~ factor(treatment), data =gerber)
exp3 <-lm(turnout04 ~ factor(treatment), data =gerber)
library(modelsummary)
modelsummary(list("Female"=exp1,"Age"=exp2,"Turnout"=exp3),coef_rename = coef_rename, title = "Balance Checks for Treatment and Control Groups",stars = TRUE)

#For explaining purposes:
#boxplot(gerber$age)
```
It works for female and turnout but not so much for age.
We know this because of the error in age. I believe this is due to lack of ceteris paribus in real world data. As the data wants us to randomly assign treatment and control groups amongst all ages evenly, we simply do not have that power with the data in the dataset. Which should be expected to a degree, as the real world rarely allows us to do such. If you run the box plot that is commented out above, you will see that the majority of our data lies within the ages of 41 and 60(roughly) with a mean of about 50. I believe this is what contributed to the error that flags us as not being random. The error would not be nearly as large if we had more of an even number across each age group. However, since we have so many data points concentrated between the ages of 41-60, its creating an error.


4. Estimate the average treatment effects of the different treatment arms whilst controlling for the variables you created for the question above. How do these estimates differ from regression estimates of the treatment effects only (i.e. without controlling for other factors)? Why?

```{r}
exp4 <-lm(voted ~ factor(treatment), data =gerber)
# Now use the same equation above but add female, age, and turnout04
exp5 <-lm(voted ~ factor(treatment) + female+age+turnout04, data=gerber)
modelsummary(list(exp4,exp5),coef_rename = coef_rename, stars = TRUE, title = "Voters actions on treatment effects only (1) and controlling for factors gender, age, and if they voted in the last election (2)")
```
When you don't control for other factors, the other factors explainability gets thrown into the error term. This causes the R squared to be lower as all factors are not being explained by the model. When we control for other factors, those factors are being taken out of the error term and help better explain our model.
The chart has treatment effects only under 1 and controlling for other factors under 2. As we pull those variables out of the error term and put them into our model, our intercept decreases significantly, as the model can now allocate what was once in our intercept to other variables that are relevant to our model. This increases our models explainability and therefore increases the R squared and the adjusted R squared (given that the variables are relevant to our model).

5. Estimate the treatment effects separately for men and women. Do you note any differences in the impact of the treatment amongst these subgroups?

```{r}
# modify the equation below for just men
exp6 <-lm(voted ~ factor(treatment), data =gerber[gerber$female== 0, ])
# modify the equation below for just women
exp7 <-lm(voted ~ factor(treatment), data =gerber[gerber$female== 1, ])
modelsummary(list("Men"=exp6,"Women"=exp7),title = "Differences in Treatment Groups Between Men and Women", coef_rename = coef_rename, star = TRUE)
```

Some notable differences between these subgroups are:
Men in the Civic Duty treatment group are more likely to vote than women in the Civic Duty treatment group. (Difference of 0.04)
Women in the Hawthorne treatment group are slightly more likely to vite than men in the Hawthorne treatment group. (Difference of 0.002)
Women in the Self treatment group are more likely to vote than men in the Self treatment group. (Difference of 0.005)
Men in the Neighbors group are slightly more likely to vote than women in the Neighbors group. (Difference of 0.001)